Fast equivalent-layer technique for gravity data processing with BTTB matrix systems

### 2019-11-29 manuscript/figures.tex
review the application to synthetic data

\plot{float}{width=\textwidth}{
	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 
	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ iterations.}
	\label{fig:float}
}

@@

\plot{synthetic_data}{width=13.5cm}{
	{Noise-corrupted gravity data (in color map) produced by three simulated sources  whose horizontal projections 
	are shown in black lines. The simulated sources are: two polygonal prisms, with density contrast of 
	$0.35\,	\mathrm{g/cm^3}$ (upper-left body) and $0.4\, \mathrm{g/cm^3}$ (upper-right body), and a sphere with 
	radius of $1\,000$ m with density contrast	of $-0.5\, \mathrm{g/cm^3}.$}
	\label{fig:synthetic_data}
}

\plot{classic_fast_val}{width=8cm}{
	{(a) Fitted gravity data produced by the fast equivalent-layer technique proposed by \citet{siqueira-etal2017}. 
	(b) Gravity residuals, defined as the difference between the observed data in Figure \ref{fig:synthetic_data} and 
	the predicted data in panel a, with their mean of $8.264e^{-7}$ and standard deviation of $0.0144$ mGal.}
	\label{fig:classic_fast_val}
}

\plot{bccb_fast_val}{width=8cm}{
	{(a) Fitted gravity data produced by our modification of the fast equivalent layer \citep{siqueira-etal2017}. 
	(b) Gravity residuals, defined as the difference between the observed data in Figure 4 and the predicted data in 
	panel a, with their mean of $8.264e^{-7}$ and standard deviation of $0.0144$ mGal.}
	\label{fig:bccb_fast_val}
}

\plot{delta_comparison}{width=13.5cm}{
	{Difference between the fitted gravity data produced by \citet{siqueira-etal2017} method 
	(Figure \ref{fig:classic_fast_val}a) and by our modified form of this method (Figure \ref{fig:bccb_fast_val}a) 
	that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}).}
	\label{fig:delta_comparison}
}

\plot{delta_rho}{width=13.5cm}{
	{Difference between the estimated mass distribution within the equivalent layer produced by \citet{siqueira-etal2017}
	method (Figure \ref{fig:classic_fast_val}) and by our modified form of this method (Figure \ref{fig:bccb_fast_val}) 
	that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}).}
	\label{fig:delta_rho}
}

\plot{upward_comparison}{width=5cm}{
	{The upward-continued gravity data using: (a) the fast equivalent layer proposed by \citet{siqueira-etal2017} and 
	(b) our modified form of \citet{siqueira-etal2017} method by using the properties of BTTB and BCCB matrices 
	(equation \ref{eq:w_Cv}) to calculate the forward modeling.  (c) Residuals, defined as the difference between 
	panels a and b with their mean of $-5.938e^{-18}$ and standard deviation of $8.701e^{-18}$.  The total computation 
	times in the \citet{siqueira-etal2017} method and in our approach are $7.62026$ and $0.00834$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{downward_comparison}{width=5cm}{
	{The downward-continued gravity data using: (a) the fast equivalent layer proposed by \citet{siqueira-etal2017} and 
	(b) our modified form of \citet{siqueira-etal2017} method by using the properties of BTTB and BCCB matrices 
	(equation \ref{eq:w_Cv}) to calculate the forward modeling.  (c) Residuals, defined as the difference between panels 
	a and b with their mean of $5.914e^{-18}$ and standard deviation of $9.014e^{-18}$.  The total computation times in 
	the \citet{siqueira-etal2017} method and in our approach are $7.59654$ and $0.00547$ seconds, respectively.}
	\label{fig:downward_comparison}
}

>>>>

% Computational performance

\plot{float}{width=\textwidth}{
	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 
	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ 
	iterations and $\kappa = 5$.}
	\label{fig:float}
}

@@

% Application to synthetic data

\plot{synthetic_data}{width=13.5cm}{
	{Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
	a sphere with density contrast $-0.5\, \mathrm{g/cm^3}$ and 
	two rectangular prisms with density contrasts $0.35\, \mathrm{g/cm^3}$ (upper-left body) and 
	$0.4\, \mathrm{g/cm^3}$ (upper-right body). The black lines represent the horizontal projection 
	of the sources.}
	\label{fig:synthetic_data}
}

\plot{classic_fast_val}{width=8cm}{
	{(a) Predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
	respectively.}
	\label{fig:classic_fast_val}
}

\plot{bccb_fast_val}{width=8cm}{
	{(a) Predicted data produced by our method. 
	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
	respectively.}
	\label{fig:bccb_fast_val}
}

\plot{delta_comparison}{width=13.5cm}{
	{Difference between the predicted data produced by the fast equivalent-layer technique 
	(Figure \ref{fig:classic_fast_val}a) and by our method (Figure \ref{fig:bccb_fast_val}a).}
	\label{fig:delta_comparison}
}

\plot{delta_rho}{width=13.5cm}{
	{Difference between the estimated mass distribution within the equivalent layers produced by 
	the fast equivalent-layer technique (not shown) and by our method (not shown).}
	\label{fig:delta_rho}
}

\plot{upward_comparison}{width=5cm}{
	{The upward-continued gravity data obtained with (a) the fast equivalent-layer technique 
	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
	panels a and b with mean $-5.938 \times 10^{-18}$ mGal and standard deviation $8.701 \times 10^{-18}$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and our method 
	are $7.62026$ and $0.00834$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{downward_comparison}{width=5cm}{
	{The downward-continued gravity data obtained with (a) the fast equivalent-layer technique 
	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
	panels a and b with mean $5.914 \times 10^{-18}$ mGal and standard deviation $9.014 \times 10^{-18}$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and our method 
	are $7.59654$ and $0.00547$ seconds, respectively.}
	\label{fig:downward_comparison}
}

% Application to field data

### 2019-11-29 manuscript/synthetic_data.tex
review the application to synthetic data

In this section, we  investigate the effectiveness of using the properties of BTTB and BCCB matrices 
(equation \ref{eq:w_Cv})  to solve, at each iteration, the forward modeling (the matrix-vector product 
$\mathbf{A} \hat{\mathbf{p}}^k$)  required in the fast equivalent-layer method proposed by \citet{siqueira-etal2017}. 
We simulated three sources whose horizontal projections are shown in Figure \ref{fig:synthetic_data} as black lines. 
These sources are two vertical prisms with density contrasts of $0.35\, \mathrm{g/cm^3}$ (upper-left prism) and 
$0.4\, \mathrm{g/cm^3}$ (upper-right prism) and a sphere with radius of $1\,000$ m with density contrast of 
$-0.5\, \mathrm{g/cm^3}$. Figure \ref{fig:synthetic_data}  shows the vertical component of gravity field generated 
by these sources contaminated with additive pseudorandom Gaussian noise with zero mean and standard deviation of 
$0.015$ mGal.

The advantage of using the structures of BTTB and BCCB matrices to compute forward modeling in  the fast-equivalent 
layer method \cite[]{siqueira-etal2017} is grounded on the use of regular grids of data and equivalent sources. 
Hence, we created $10\,000$ observation points regularly spaced in a grid of $100 \times 100$ at $100$ m height. 
We also set a grid of equivalent point masses, each one directly beneath each observation points, located at 
$300$ m deep.  Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a  show the fitted gravity data 
obtained, respectively, by the fast equivalent layer method and by our modified form of this method that computes 
the forward modeling using equation \ref{eq:w_Cv}. The corresponding residuals (Figures \ref{fig:classic_fast_val}b 
and \ref{fig:bccb_fast_val}b), defined as the difference between the observed (Figure  \ref{fig:synthetic_data}) 
and fitted gravity data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a), show means close to 
zero  and standard deviations of 0.0144 mGal.  Therefore, Figures \ref{fig:classic_fast_val} and 
\ref{fig:bccb_fast_val} show that \citet{siqueira-etal2017} method and our modified version of this method produced 
virtually the same results. This excellent agreement is confirmed in Figures \ref{fig:delta_comparison} and 
\ref{fig:delta_rho} which shows that there are virtually no differences, respectively,  in the fitted data presented 
in Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b and in the estimated mass distributions within 
the equivalent layers (not shown) yielded by both \citet{siqueira-etal2017} method and our modification of this method. 
These results (Figures \ref{fig:delta_comparison} and \ref{fig:delta_rho})  show that computing the matrix-vector 
product ($\mathbf{A} \hat{\mathbf{p}}^k$), required in the forward modeling, by means of embedding the BTTB matrix into 
a BCCB matrix (equation \ref{eq:w_Cv}) yields practically the same result as the one produced by computing this 
matrix-vector product  with a full matrix $\mathbf{A}$ as used in \citet{siqueira-etal2017}.

We perform two forms of processing the gravity data (Figure \ref{fig:synthetic_data}) through the equivalent layer 
technique: the upward (Figure \ref{fig:upward_comparison}) and the downward (Figure \ref{fig:downward_comparison}) 
continuations. The upward height is 300 m and the downward  is at 50 m.   Either in the upward continuation 
(Figure \ref{fig:upward_comparison}) or in the downward continuation (Figure \ref{fig:downward_comparison}), the 
continued gravity data using the fast equivalent layer proposed by \citet{siqueira-etal2017} 
(Figures \ref{fig:upward_comparison}a and \ref{fig:downward_comparison}a) are in close agreement with those produced 
by our modification of \citet{siqueira-etal2017} method (Figures \ref{fig:upward_comparison}b and 
\ref{fig:downward_comparison}b). The residuals (Figures \ref{fig:upward_comparison}c and \ref{fig:downward_comparison}c)
quantify this agreement since their means and standard deviations are close to zero in both continued gravity data using 
both methods.  All the continued gravity data shown here (Figures \ref{fig:upward_comparison} and 
\ref{fig:downward_comparison}) agree with the true ones (not shown). The most striking feature of these upward or 
the downward continuations concerns the total computation time. The computation time spent by our method is 
approximately $1\,500$ times faster than \citet{siqueira-etal2017} method.

>>>>

We simulated three sources whose horizontal projections are shown in Figure \ref{fig:synthetic_data} as black lines. 
These sources are a sphere with radius of $1\,000$ m with density contrast of 
$-0.5\, \mathrm{g/cm^3}$ and two rectangular prisms with density contrasts of $0.35\, \mathrm{g/cm^3}$ (upper-left prism) 
and $0.4\, \mathrm{g/cm^3}$ (upper-right prism). 
Figure \ref{fig:synthetic_data} shows the gravity disturbance (vertical component of gravitational attraction) 
produced by these sources. The synthetic data are contaminated with additive pseudo-random Gaussian 
noise with zero mean and standard deviation of $0.015$ mGal.
We created $10\,000$ observation points regularly spaced on a grid of $100 \times 100$, on a 
horizontal plane at $-100$ m, and set a grid of equivalent point masses, each one directly beneath each 
observation point, located at depth $300$ m.

Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a show the data fits 
obtained, respectively, by the fast equivalent-layer technique \citep{siqueira-etal2017} 
and by our method. The corresponding residuals (Figures \ref{fig:classic_fast_val}b 
and \ref{fig:bccb_fast_val}b), defined as the difference between the simulated (Figure \ref{fig:synthetic_data}) 
and predicted data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a), show means close to 
zero and standard deviations of $0.0144$ mGal, indicating that both methods produced virtually the same results.
This excellent agreement is confirmed by Figures \ref{fig:delta_comparison} and 
\ref{fig:delta_rho}. They show that there are virtually no differences between, respectively, 
the predicted data (Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b) and the 
estimated mass distributions within the equivalent layers (not shown) yielded by 
both methods.

We performed the upward and downward continuations of the simulated gravity data 
(Figure \ref{fig:synthetic_data}) by using the estimated equivalent layers.
The results show that the continued gravity data obtained by using the 
fast equivalent-layer (Figures \ref{fig:upward_comparison}a and \ref{fig:downward_comparison}a) 
are in close agreement with those produced by our method (Figures \ref{fig:upward_comparison}b and 
\ref{fig:downward_comparison}b). The residuals (Figures \ref{fig:upward_comparison}c and \ref{fig:downward_comparison}c)
quantify this agreement since their means and standard deviations are close to zero.

The continued gravity data obtained with both methods agree with the true ones (not shown).
The difference is that the computation time spent by our method is 
approximately $1\,500$ times faster than that of the fast-equivalent-layer 
technique \citep{siqueira-etal2017}.
